---
title: "Train model for Human PBMC"
author: "Christoph Hafemeister"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: pygments
    df_print: kable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  digits = 2,
  tidy = TRUE,
  tidy.opts = list(width.cutoff=80)
)
options(tibble.width = Inf)
set.seed(6646428)
```

## Introduction

We use the Human PBMC [Azimuth reference](https://azimuth.hubmapconsortium.org/references/#Human%20-%20PBMC) of 161,764 cells grouped into 30 types (annotation level 2). 


[Azimuth](https://azimuth.hubmapconsortium.org/) is a [Seurat](https://satijalab.org/seurat/) workflow that leverages high-quality references to automate visualization and annotation of scRNA-seq data. See [Hao and Hao et al, bioRvix 2020](https://satijalab.org/v4preprint) for details.

## Load the Azimuth human PBMC data to use as training data

```{r}
library('Matrix')

s_multi <- SeuratDisk::Connect(filename = '~/Downloads/pbmc_multimodal.h5seurat')
counts <- Seurat::as.sparse(s_multi[["assays/SCT/counts"]])
features <- as.character(as.array(s_multi[["assays/SCT/features"]]))
meta_data <- as.data.frame(s_multi[["meta.data"]])
dimnames(counts) <- list(features, rownames(meta_data))
celltype <- factor(meta_data$celltype.l2)
```

## Load pre-calculated DE results

The DE test was run with the following commands (took ca. 3 hours)
```{r, eval=FALSE}
de_res <- sctransform::diff_mean_test(y = counts, group_labels = celltype, 
                                      R = 499, log2FC_th = 0, mean_th = 0.05,
                                      compare = 'all_vs_all', verbosity = 1)
saveRDS(de_res, file = '~/Downloads/pbmc_multimodal_de_res_all_vs_all.rds')
```

```{r}
de_res <- readRDS(file = '~/Downloads/pbmc_multimodal_de_res_all_vs_all.rds')
```

## Train the model

```{r}
pg_model_human_pbmc <- Perigee::pg_train(
  counts = counts, cell_labels = celltype, de_res = de_res,
  max_cells = 5000, exclude = 'Doublet', verbosity = 2)
```

## Save the model

```{r}
save(pg_model_human_pbmc, file = '../data/pg_model_human_pbmc.rda')
```
