---
title: "Intro to Perigee"
author: "Christoph Hafemeister"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: pygments
    df_print: kable
---

```{r setup, include = FALSE}
library('ggplot2')
library('knitr')
knit_hooks$set(optipng = hook_optipng)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  digits = 2,
  tidy = TRUE,
  tidy.opts = list(width.cutoff=80),
  optipng = '-o 5 -strip all -quiet',
  fig.width=8, fig.height=4, dpi=100, out.width = '90%'
)
old_theme <- theme_set(theme_bw(base_size=10))
options(tibble.width = Inf)
set.seed(6646428)
```

## Introduction

The Perigee package provides training and prediction methods for cell type classification of single-cell data. Some pre-trained models come as part of the package.

Perigee classifies every cell on its own, i.e. the context (composition of the dataset) is irrelevant and even a single cell can be classified. Cell type prediction is fast, taking about 6 seconds for 11k cells and the 30 cell type human PBMC model.

## PBMC cell type prediction

As a first example we will show how to use a pre-trained model to predict cell types.

As example data we load the public 10x Genomics 10k PBMC v3 raw counts (11,769 cells)
```{r}
counts <- Seurat::Read10X_h5(filename = '~/Projects/data_warehouse/raw_public_10x/pbmc_10k_v3_filtered_feature_bc_matrix.h5')
```

Type prediction (aka label transfer) is done using the `pg_predict` function. The `pg_model_human_pbmc` object is provided with the package.
```{r}
library('Perigee')
prediction <- pg_predict(pg_model = pg_model_human_pbmc, counts = counts)
head(prediction)
as.data.frame(table(prediction$label, useNA = 'always'))
```

## Compare to Azimuth

The data used above is the same as the demo dataset used in [Azimuth](https://app.azimuth.hubmapconsortium.org/app/human-pbmc). How do the predictions compare? 

Load the results obtained from Azimuth (Azimuth version: 0.3.2; Seurat version: 4.0.0; Reference version: 1.0.0; default parameters; web-app took 82 seconds for visualization, label transfer, protein signal transfer, DE testing).
```{r}
azimuth_pred <- read.delim('~/Projects/data_warehouse/raw_public_10x/pbmc_10k_v3_azimuth_pred.tsv', row.names = 1)
# make sure predictions are in the same order
identical(rownames(prediction), rownames(azimuth_pred))
head(azimuth_pred)
as.data.frame(table(azimuth_pred$predicted.celltype.l2))
```

For each cell type, get Jaccard similarity coefficient of the two predictions (intersection over union of cells predicted to be of a given type)
```{r}
cell_types <- union(prediction$label, azimuth_pred$predicted.celltype.l2)
cell_types <- cell_types[!is.na(cell_types)]
jsc <- sapply(cell_types, function(type) {
  tab <- table(factor(prediction$label == type, levels = c(FALSE, TRUE)),
               factor(azimuth_pred$predicted.celltype.l2 == type, levels = c(FALSE, TRUE)))
  tab[2,2] / (tab[1,2] + tab[2,2] + tab[2,1])
})
sort(jsc, decreasing = TRUE)
```

Visually compare the predictions using the Azimuth UMAP coordinates.
```{r}
azimuth_umap <- readRDS('~/Projects/data_warehouse/raw_public_10x/pbmc_10k_v3_azimuth_umap.Rds')
df <- cbind(prediction, azimuth_pred, azimuth_umap@cell.embeddings)
df$label <- factor(df$label, levels = sort(cell_types))
df$predicted.celltype.l2 <- factor(df$predicted.celltype.l2, levels = sort(cell_types))
library('ggplot2')
ggplot(df, aes(UMAP_1, UMAP_2, color = label)) + geom_point()
ggplot(df, aes(UMAP_1, UMAP_2, color = predicted.celltype.l2)) + geom_point()
```

The cells where the two methods disagree tend to have a lower prediction score in Azimuth
```{r}
ggplot(df, aes(label == predicted.celltype.l2, predicted.celltype.l2.score)) + geom_violin()
```

TODO: Look in depth at prediction differences.


## Session info 

Session info
```{r}
sessionInfo()
```
